---
title: "Text Analysis of Science books"
author: "Jajwalya R. Karajgikar"  
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: embed
---
#**************************************************************************
#                  Analyzing a corpus of Science books
#**************************************************************************
#set working directory 
#setwd("replace directory here")
#install and load the dplyr, tidytext, ggplot2, and gutenbergr library for text processing
```{r}
#install.packages("dplyr")
#install.packages("tidytext")
#install.packages("ggplot2")
#install.packages("gutenbergr")
```
#**************************************************************************
#                  Creating a corpus
#**************************************************************************
```{r}
library(dplyr)
library(tidytext)
library(ggplot2)
library(gutenbergr)
physics <- gutenberg_download(c(37729, 14725, 13476, 30155), 
                              meta_fields = "author")
```

#How many times was each word used in each text?

```{r}
data(stop_words)
physics_words <- physics %>%
  unnest_tokens(word, text) %>%
  count(author, word, sort = TRUE) %>%
  anti_join(stop_words)

physics_words
```



```{r}
library(forcats)

plot_physics <- physics_words %>%
  bind_tf_idf(word, author, n) %>%
  mutate(word = fct_reorder(word, tf_idf)) %>%
  mutate(author = factor(author, levels = c("Galilei, Galileo",
                                            "Huygens, Christiaan", 
                                            "Tesla, Nikola",
                                            "Einstein, Albert")))
```

Row {data-height=600}
------------------------------------------------------------------------------

### Chart 1
```{r}

plot_physics %>% 
  group_by(author) %>% 
  top_n(15, tf_idf) %>% 
  ungroup() %>%
  mutate(word = reorder(word, tf_idf)) %>%
  ggplot(aes(word, tf_idf, fill = author)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~author, ncol = 2, scales = "free") +
  coord_flip()
```




```{r}
library(stringr)

physics %>% 
  filter(str_detect(text, "_k_")) %>% 
  select(text)

physics %>% 
  filter(str_detect(text, "RC")) %>% 
  select(text)

mystopwords <- tibble(word = c("eq", "co", "rc", "ac", "ak", "bn", 
                               "fig", "file", "cg", "cb", "cm",
                               "ab", "_k", "_k_", "_x"))

physics_words <- anti_join(physics_words, mystopwords, 
                           by = "word")
head(physics_words)
```

```{r}
library("SnowballC")
library("wordcloud")
library("RColorBrewer")

physics_words %>%
  count(word) 

wordcloud(physics_words$word, physics_words$n, min.freq = 50, max.words = 200, random.order=FALSE, rot.per=0.35, colors=brewer.pal(8, "Dark2"))
      
```
Row {data-height=400}
------------------------------------------------------------------------------

### Chart 2
```{r}
plot_physics <- physics_words %>%
  bind_tf_idf(word, author, n) %>%
  mutate(word = str_remove_all(word, "_")) %>%
  group_by(author) %>% 
  top_n(15, tf_idf) %>%
  ungroup() %>%
  mutate(word = reorder_within(word, tf_idf, author)) %>%
  mutate(author = factor(author, levels = c("Galilei, Galileo",
                                            "Huygens, Christiaan",
                                            "Tesla, Nikola",
                                            "Einstein, Albert")))

ggplot(plot_physics, aes(word, tf_idf, fill = author)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~author, ncol = 2, scales = "free") +
  coord_flip() +
  scale_x_reordered()
```

```{r}
library(tidyr)

frequency <- bind_rows(mutate(plot_physics, author = "Galilei, Galileo"),
                       mutate(plot_physics, author = "Tesla, Nikola"), 
                       mutate(plot_physics, author = "Einstein, Albert")) %>% 
  mutate(word = str_extract(word, "[a-z']+")) %>%
  count(author, word) %>%
  group_by(author) %>%
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  spread(author, proportion) %>% 
  gather(author, proportion, `Galilei, Galileo`:`Tesla, Nikola`)
```

```{r}
library(scales)

# expect a warning about rows with missing values being removed
ggplot(frequency, aes(x = proportion, y = `Einstein, Albert`, color = abs(`Einstein, Albert` - proportion))) +
  geom_abline(color = "gray40", lty = 2) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
  facet_wrap(~author, ncol = 2) +
  theme(legend.position="none") +
  labs(y = "Einstein, Albert", x = NULL)
```

```{r}
physics_bigrams <- physics %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2)

physics_bigrams
```

```{r}
physics_bigrams%>%
  count(bigram, sort = TRUE)
```

```{r}
library(tidyr)

bigrams_separated <- physics_bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

# new bigram counts:
bigram_counts <- bigrams_filtered %>% 
  count(word1, word2, sort = TRUE)

bigram_counts
```

```{r}
bigrams_united <- bigrams_filtered %>%
  unite(bigram, word1, word2, sep = " ")

bigrams_united
```

```{r}
physics %>%
  unnest_tokens(trigram, text, token = "ngrams", n = 3) %>%
  separate(trigram, c("word1", "word2", "word3"), sep = " ") %>%
  filter(!word1 %in% stop_words$word,
         !word2 %in% stop_words$word,
         !word3 %in% stop_words$word) %>%
  count(word1, word2, word3, sort = TRUE)
```

```{r}
bigrams_filtered %>%
  filter(word2 == "angle") %>%
  count(author, word1, sort = TRUE)
```
Row {data-height=600}
------------------------------------------------------------------------------

### Chart 3

```{r}
bigram_tf_idf <- bigrams_united %>%
  count(author, bigram) %>%
  bind_tf_idf(bigram, author, n) %>%
  arrange(desc(tf_idf))

bigram_tf_idf
```
```{r}
library(igraph)

bigram_graph <- bigram_counts %>%
  filter(n > 20) %>%
  graph_from_data_frame()

bigram_graph
```
Row {data-height=600}
------------------------------------------------------------------------------

### Chart 4

```{r}
library(ggraph)
set.seed(2000)

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1)
```

Row {data-height=600}
------------------------------------------------------------------------------

### Chart 5 

```{r}
set.seed(2000)

a <- grid::arrow(type = "closed", length = unit(.15, "inches"))

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                 arrow = a, end_cap = circle(.07, 'inches')) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()
```


#**************************************************************************
#                    Citations
#**************************************************************************

This project is heavily adapted from the [website](https://www.tidytextmining.com/) for Text Mining with R! Visit the [GitHub repository](https://github.com/dgrtwo/tidy-text-mining) for this site, find the [book at O’Reilly](http://shop.oreilly.com/product/0636920067153.do?cmp=af-strata-books-video-product_cj_0636920067153_4428796), or buy it on [https://www.amazon.com/gp/product/1491981652/ref=as_li_tl?ie=UTF8&tag=juliasilge-20&camp=1789&creative=9325&linkCode=as2&creativeASIN=1491981652&linkId=0e92d44b0aa39ab34608ffa582dbd490](Amazon).


citation("dplyr")
citation("tidytext")
citation("ggplot2")
citation("gutenbergr")
citation("forcats")
citation("stringr")
citation("wordcloud")
citation("tidyr")
citation("igraph")
citation("ggraph")

#Silge J. and Robinson D. (2017) Text Mining with R: A Tidy Approach (1st. ed.). O’Reilly Media, Inc.